---
title: 'Reproducible Research: Peer Assessment 1'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data

```{r loading data}
# Loading libraries
library(dplyr)
library(ggplot2)
# Reading data
data <- read.csv("activity.csv")
# Setting the date format
data$date <- as.Date(data$date, format = "%Y-%m-%d")
# Working out total number of steps per day
totsteps <- group_by(data, date)
totsteps <- summarize(totsteps, sum(steps))
names(totsteps) <- c("date", "totalsteps")
# Summarizing by Intervals
avgint <- group_by(data, interval)
avgint <- summarize(avgint, mean(steps, na.rm = TRUE))
names(avgint) <- c("interval", "avgsteps")

```


## What is mean total number of steps taken per day?

```{r meantotsteps}

qplot(totsteps$totalsteps, data = totsteps, main = "Total Number of Steps Taken Each Day", geom = "histogram")
totstepmean <- round(mean(totsteps$totalsteps, na.rm = TRUE), digits = 2)
totstepmed <- median(totsteps$totalsteps, na.rm = TRUE)

```



The **Mean** total number of steps each day is **`r totstepmean `** and the **Median** total number of steps each day is **`r totstepmed`**.


## What is the average daily activity pattern?

```{r dailyact}
qplot(avgint$interval, avgint$avgsteps, data = avgint, main = "Daily Activity Pattern", geom = "line")
maxsteps <- avgint[which(avgint$avgsteps==max(avgint$avgsteps)),]

```

As we can see from the output, the 5-minute interval with the maximum number is steps is interval = `r maxsteps$interval`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?




